USAMS Hybrid Source Summary

### 2021-01-19 hgis_summary.Rmd
Update summary.

Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite
currents. This is in line with what other hybrid sources produce.  The source
typically produces best current at around 10uLs-1 CO2. Literature values give
1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
flows in at least one of these papers doesn't make sense.  Precision on modern
targets has been as good as 1% (SD of multiple measurements of a target) with
directly introduced bottle gas. Typical precision in tests of inlet systems has
been around 3%. Measured value of tank modern gas agrees with the accepted
value within error when normalizing to solid graphite standards.  The system
seems to have a constant blank contribution from the target and source of about
100nA 12C- or 1% of a sample producing 10uA 12C-. Helium does not seem to add
significantly to this blank. Since the blank current is constant, the blank
contribution is inversely related to the sample current, ie. if the sample
current is 5uA, 2% of that will be from the blank and so on.

@@

The open split interface supplies a helium-co2 mixture to a split where a
capillary delivers ~10uL/min to the source and the rest is vented. The current
split starts showing "breakthrough" of lab air at around 150uLs-1 flow.
Currently using 3m of 50um capillary with a bulkhead connector at the source
cage.  Precision of measurements of modern CO2 has been 3% at best, 10% at
worst, with better precision most recently. This is much less precise than the
expected 1%.  Accuracy is reasonable within the large errors on gas
measurements. Uncorrected measurements of dead CO2 range from 1% to 5%
depending on sample current.

@@

Q: If the relative blank contribution is proportional to sample current, can we
correct for a varying sample current over a run (due to variable dillution,
pressure, etc.)?

A: If the sample current varies widely, this will likely require taking data in
shorter segments and blank correcting each segment using the mean current for
that segment.

@@

A: Unknown. We can rerun a test with the best performing conditions (bottle gas
through capillary to source) to eliminate as many variables as possible.

## Development

>>>>

Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite
currents. This is in line with what other hybrid sources produce.  The source
typically produces best current at around 10uLs-1 CO2. Literature values give
1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
flows in at least one of these papers doesn't make sense.

Initially, precision on modern targets was thought to be as good as 1% (SD of
multiple measurements of a modern target) with directly introduced bottle gas,
which is in agreement with literature values. Revisiting these data show
precision in the range of 2-4%. Typical precision in tests of inlet systems has
been in the range of 2-4%.

Measured value of tank modern gas agrees with the accepted value within error
when normalizing to solid graphite standards.  The system seems to have a
constant blank contribution from the target and source of about 100nA 12C- or 1%
of a sample producing 10uA 12C-. Helium does not seem to add significantly to
this blank. Since the blank current is constant, the blank contribution is
inversely related to the sample current, ie. if the sample current is 5uA, 2% of
that will be from the blank and so on.

The above measurements are at optimum sample current of 6-10uA 12C-. Precision
and measured ratio both decrease as sample current drops below the optimum
range. The ratio effects may be related to the "constant current" blank
discussed above.

@@

The open split interface supplies a helium-co2 mixture to a split where a
capillary delivers ~10uL/min to the source and the rest is vented. The current
split starts showing "breakthrough" of lab air at around 150uL/min flow.
Currently using 3m of 50um capillary with a bulkhead connector at the source
cage.  Precision of measurements of modern CO2 has been 2% at best, 10% at
worst, with better precision most recently. This is much less precise than the
expected 1%.  Accuracy is reasonable within the large errors on gas
measurements. Uncorrected measurements of dead CO2 range from 1% to 5%
depending on sample current.

## Helium displacement from vials

First tests of displacing pure CO2 from a vial into the open split using helium
are promising. Helium flowing at 100uLm-1 produces current from 7mL of CO2 for
over 2 hours. Currents do decrease with time due to dilution, but by supplying
higher than optimal CO2 flow to the source initially, stable currents can be
maintained for some time.

@@

Q: If the relative blank contribution is proportional to sample current, can we
correct for a varying sample current over a run (due to variable dilution,
pressure, etc.)?

A: If the sample current varies widely, this will likely require taking data in
shorter segments and blank correcting each segment using the mean current for
that segment. We may be able to mitigate by setting capillary length such that
CO2 flow starts above optimal and drops to optimal by the end of the
acquisition.

@@

A: Unknown. Tests with the best performing conditions (bottle gas through
capillary to source) give 2% precision, and revisiting early test data shows
precision of 2%. Precision is driven by variability between measurements of a
sample. Investigating correlations here may help. We should also look at
variability between identical samples to get a sense of reproducibility.

## Development
2021-01-15 - Pure CO2 and displacement tests using vial and double needle. Tested precision with vial setup by flowing CO2 at 200kPa (high flow, roughly 400ul/min) into open split. Followed this by switching to helium flow to vial at 100kPa (200ul/min). Repeated test with a second vial. Built second dual needle setup for use in vial filling (can also use Gilson).

2021-01-08 - Precision tests with direct CO2-source connection. Tested source pressure and current vs CO2 regulator pressure. Tested flow vs pressure with leak checker. Several capillary swaps. Acquired data at several flow rates for precision testing.

2021-01-07 - Set up for precision test using 7m x 50um capillary from CO2/he via Valco valve to cage wall.

2020-12-31 - Add CO2/He valve to vial test setup. VICI 4-way valve connected
with 1/16" ss tubing from CO2 and He regulators allows switching between gasses
delivered to a glass capillary leading to outer needle of dual needle. Leak
checking and capillary adjustments for good flow range at reasonable pressures.

2020-12-24 - Build dual needle setup. Supply displacement helium through outer
1/16" tubing, sample mixture through inner needle. Use VICI Luer adapter on top
of needle as open split with inserted capillary to source.
